---
title: "Understanding the Regional Differences in World Happiness Index"
output: html_document
date: "2023-04-05"
---

---

Institution: Chaitanya Bharathi Institute of Technology

Branch: AI & DS

Students: Mayush T - 160120771039, Rithvik R - 160120771043

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(cache = TRUE, results="markup", eval = FALSE)
```


## Loading Libraries
```{r}
library(caret)
library(readr)
library(dplyr)
library(ggplot2)
library(skimr)
library(tidyr)
library(ggpubr)
```

## Loading the data
```{r, eval=FALSE}
df_15 = read.csv("2015.csv")
df_16 = read.csv("2016.csv")
df_17 = read.csv("2017.csv")
df_18 = read.csv("2018.csv")
df_19 = read.csv("2019.csv")
df_20 = read.csv("2020.csv")
df_21 = read.csv("2021.csv")
df_22 = read.csv("2022.csv")
head(df_19)
```
## Renaming dataset column names

### Changing year 2015 datasets column names based year 2017 dataset
```{r}
df_15=plyr::rename(df_15, replace = c("Happiness Rank" = "Happiness.Rank", 
                                  "Happiness Score" = "Happiness.Score",
                                  "Economy (GDP per Capita)" = "Economy..GDP.per.Capita.",
                                  "Health (Life Expectancy)" = "Health..Life.Expectancy.",
                                  "Trust (Government Corruption)" = "Trust..Government.Corruption.",
                                  "Dystopia Residual"="Dystopia.Residual"
                                  ))
str(df_15)
```

### Changing year 2016 datasets column names based on year 2017 dataset
```{r}
df_16=plyr::rename(df_16, replace = c( "Happiness Rank" = "Happiness.Rank", 
                                  "Happiness Score" = "Happiness.Score",
                                  "Economy (GDP per Capita)" = "Economy..GDP.per.Capita.",
                                  "Health (Life Expectancy)" = "Health..Life.Expectancy.",
                                  "Trust (Government Corruption)"  = "Trust..Government.Corruption.",
                                  "Dystopia Residual"="Dystopia.Residual"
                                  ))
str(df_16)
```

### Changing year 2018 datasets column names based on year 2017 dataset
```{r}
df_18=plyr::rename(df_18, replace = c( "Country.or.region"="Country", 
                                  "Overall.rank"="Happiness.Rank" ,
                                  "GDP.per.capita"="Economy..GDP.per.Capita.",
                                  "Healthy.life.expectancy"="Health..Life.Expectancy.",
                                  "Freedom.to.make.life.choices"="Freedom",
                                  "Perceptions.of.corruption"="Trust..Government.Corruption.",
                                  "Social.support"="Family",
                                  "Score"="Happiness.Score"))
str(df_18)
```

### Changing year 2019 datasets column names based on year 2017 dataset
```{r}
df_19=plyr::rename(df_19, replace = c( "Country.or.region"="Country", 
                                  "Overall.rank"="Happiness.Rank" ,
                                  "GDP.per.capita"="Economy..GDP.per.Capita.",
                                  "Healthy.life.expectancy"="Health..Life.Expectancy.",
                                  "Freedom.to.make.life.choices"="Freedom",
                                  "Perceptions.of.corruption"="Trust..Government.Corruption.",
                                  "Social.support"="Family",
                                  "Score"="Happiness.Score"))
str(df_19)
```
### Changing year 2020 datasets column names based on year 2017 dataset
```{r}
# Adding a happiness rank column as it is not present in the dataset
df_20$Happiness.Rank <- rank(-df_20$Ladder.score)

# Arranging columns
df_20 <- df_20 %>% 
  select(Country.name, Happiness.Rank, everything())

# Renaming the necessary column names
df_20 = plyr::rename(df_20, replace = c(
  "Country.name"="Country", 
  "Logged.GDP.per.capita"="Economy..GDP.per.Capita.",
  "Healthy.life.expectancy"="Health..Life.Expectancy.",
  "Freedom.to.make.life.choices"="Freedom",
  "Perceptions.of.corruption"="Trust..Government.Corruption.",
  "Social.support"="Family",
  "Ladder.score"="Happiness.Score"
))

# Convert to numeric
df_20$Happiness.Score <- as.numeric(df_20$Happiness.Score)
df_20$Economy..GDP.per.Capita. <- as.numeric(df_20$Economy..GDP.per.Capita.)/10
df_20$Health..Life.Expectancy. <- as.numeric(df_20$Health..Life.Expectancy.)/100
str(df_20)
```

### Changing year 2021 datasets column names based on year 2017 dataset
```{r}
# Adding a happiness rank column as it is not present in the dataset
df_21$Happiness.Rank <- rank(-df_21$Ladder.score)

df_21 = plyr::rename(df_21, replace = c(
  "Country.name"="Country", 
  "Logged.GDP.per.capita"="Economy..GDP.per.Capita.",
  "Healthy.life.expectancy"="Health..Life.Expectancy.",
  "Freedom.to.make.life.choices"="Freedom",
  "Perceptions.of.corruption"="Trust..Government.Corruption.",
  "Social.support"="Family",
  "Ladder.score"="Happiness.Score"
))

# Arranging columns
df_21 <- df_21 %>% 
  select(Country, Happiness.Rank, everything())

df_21$Economy..GDP.per.Capita. <- as.numeric(df_21$Economy..GDP.per.Capita.)/10
df_21$Health..Life.Expectancy. <- as.numeric(df_21$Health..Life.Expectancy.)/100

str(df_21)
```

### Changing year 2022 dataset column names based on year 2017 dataset
```{r}
df_22 = plyr::rename(df_22, replace = c(
  "RANK"="Happiness.Rank" ,
  "Explained.by..GDP.per.capita"="Economy..GDP.per.Capita.",
  "Explained.by..Healthy.life.expectancy"="Health..Life.Expectancy.",
  "Explained.by..Freedom.to.make.life.choices"="Freedom",
  "Explained.by..Perceptions.of.corruption"="Trust..Government.Corruption.",
  "Explained.by..Social.support"="Family",
  "Score"="Happiness.Score"
))
df_22$Happiness.Score <- as.numeric(gsub(",", "", df_22$Happiness.score))/1000
df_22$Whisker.high <- as.numeric(gsub(",", ".", df_22$Whisker.high))
df_22$Whisker.low <- as.numeric(gsub(",", ".", df_22$Whisker.low))
df_22$Economy..GDP.per.Capita. <- as.numeric(gsub(",", ".", df_22$Economy..GDP.per.Capita.))
df_22$Family <- as.numeric(gsub(",", ".", df_22$Family))
df_22$Health..Life.Expectancy. <- as.numeric(gsub(",", ".", df_22$Health..Life.Expectancy.))
df_22$Freedom <- as.numeric(gsub(",", ".", df_22$Freedom))
df_22$Explained.by..Generosity <- as.numeric(gsub(",", ".", df_22$Explained.by..Generosity))
df_22$Generosity <- as.numeric(gsub(",", ".", df_22$Explained.by..Generosity))
df_22$Trust..Government.Corruption. <- as.numeric(gsub(",", ".", df_22$Trust..Government.Corruption.))
df_22$Dystopia..1.83....residual <- as.numeric(gsub(",", ".", df_22$Dystopia..1.83....residual))
str(df_22)
colnames(df_22)
```


### Insert year column at first position (index 0)
```{r}
df_15<-cbind(Year=2015,df_15)

df_16<-cbind(Year=2016,df_16)

df_17<-cbind(Year=2017,df_17)

df_18<-cbind(Year=2018,df_18)

df_19<-cbind(Year=2019,df_19)

df_20<-cbind(Year=2020,df_20)

df_21<-cbind(Year=2021,df_21)

df_22<-cbind(Year=2022,df_22)
```

### Change column type for emerging dataset
```{r}
df_18$Trust..Government.Corruption. = as.numeric(df_18$Trust..Government.Corruption.)

str(df_18)
```


### Merge data from 2015-2022
```{r}
df15_16<-dplyr::bind_rows(df_15,df_16)

df15_16_17<-dplyr::bind_rows(df15_16,df_17)

df18_19<-dplyr::bind_rows(df_18,df_19)

df_20_21 <- dplyr::bind_rows(df_20,df_21)

str(df_20_21)

df_20_21_22 <- dplyr::bind_rows(df_20_21,df_22)

df_15_16_17_18_19<-dplyr::bind_rows(df18_19,df15_16_17)

df<-dplyr::bind_rows(df_15_16_17_18_19, df_20_21_22)

head(df)

```



## Change Happiness.Rank data type
```{r}
df$Happiness.Rank  = as.numeric(df$Happiness.Rank)

df <- df %>% 
  select(Year, Happiness.Rank, Country, Region, Happiness.Score, Economy..GDP.per.Capita., Family, Health..Life.Expectancy., Freedom, Generosity, Trust..Government.Corruption.)

str(df)
```
Remove unnecessary Columns

```{r}
colSums(is.na(df))
```
## Impute with mean or median values for numerical column
```{r}
df$Trust..Government.Corruption.[is.na(df$Trust..Government.Corruption.)] <- median(df$Trust..Government.Corruption., na.rm = T)

df$Happiness.Score[is.na(df$Happiness.Score)] <- median(df$Happiness.Score, na.rm = T)

df$Economy..GDP.per.Capita.[is.na(df$Economy..GDP.per.Capita.)] <- median(df$Economy..GDP.per.Capita., na.rm = T)

df$Family[is.na(df$Family)] <- median(df$Family, na.rm = T)

df$Health..Life.Expectancy.[is.na(df$Health..Life.Expectancy.)] <- median(df$Health..Life.Expectancy., na.rm = T)

df$Freedom[is.na(df$Freedom)] <- median(df$Freedom, na.rm = T)

df$Generosity[is.na(df$Generosity)] <- median(df$Generosity, na.rm = T)

colSums(is.na(df))
```
## Filter uncommon data in Country Column

```{r}
aggregate(df$Country, by=list(df$Year), FUN=length)
```
```{r}
Country_2015 = subset(df, Year == 2015)$Country
Country_2016 = subset(df, Year == 2016)$Country
Country_2017 = subset(df, Year == 2017)$Country
Country_2018 = subset(df, Year == 2018)$Country
Country_2019 = subset(df, Year == 2019)$Country
Country_2020 = subset(df, Year == 2020)$Country
Country_2021 = subset(df, Year == 2021)$Country
Country_2022 = subset(df, Year == 2022)$Country
```


```{r}
common_country =intersect(intersect(intersect(intersect(intersect(intersect(intersect(Country_2015,
Country_2016),Country_2017),Country_2018),Country_2019),Country_2020),Country_2021),Country_2022)
length(common_country)
```
```{r}
df1 = subset(df,Country %in% common_country)
print(paste("The amount of rows in the dataset is: ",dim(df1)[1]))
print(paste("The amount of columns in the dataset is: ",dim(df1)[2]))
```
## Fill value for categorical columns
```{r}
common_region <- unique(subset(df1, Region!="NA", c(Country, Region)))

head(common_country)
```

```{r}
assign_region <- function(x){
  Region <- common_region$Region[common_region$Country == x]
}

for(country in common_country)
      df1$Region[df1$Country == country] <- assign_region(country)
```

## Saving cleaned dataset
```{r}
write_csv(df1, path = "15-22_cleaned.csv")
```

```{r}
skimr::skim_without_charts(df1)
```
## Exploratory Data analysis

### Explore data by country, region and year

#### Top 10 happiest countries
##### Top 10 happiest countries in 2015

```{r}
df1 %>%
  filter(Year == 2015) %>%
  arrange(-Happiness.Score) %>%
  slice_head(n=10) %>%
  ggplot(aes(reorder(Country, Happiness.Score), Happiness.Score)) +
  geom_point(colour = "blue", size = 3) +
  theme(text=element_text(size=10)) + 
  coord_flip() +
  labs(title = "The 10 happiest countries in 2015", x = "")
```

##### Top 10 happiest countries in 2016

```{r}
df1 %>%
  filter(Year == 2016) %>%
  arrange(-Happiness.Score) %>%
  slice_head(n=10) %>%
  ggplot(aes(reorder(Country, Happiness.Score), Happiness.Score)) +
  geom_point(colour = "blue", size = 3) +
  theme(text=element_text(size=10)) + 
  coord_flip() +
  labs(title = "The 10 happiest countries in 2016", x = "")
```

#### Top 10 happiest countries in 2017

```{r}
df1 %>%
  filter(Year == 2017) %>%
  arrange(-Happiness.Score) %>%
  slice_head(n=10) %>%
  ggplot(aes(reorder(Country, Happiness.Score), Happiness.Score)) +
  geom_point(colour = "blue", size = 3) +
  theme(text=element_text(size=10)) + 
  coord_flip() +
  labs(title = "The 10 happiest countries in 2017", x = "")
```

#### Top 10 happiest countries in 2018

```{r}
df1 %>%
  filter(Year == 2018) %>%
  arrange(-Happiness.Score) %>%
  slice_head(n=10) %>%
  ggplot(aes(reorder(Country, Happiness.Score), Happiness.Score)) +
  geom_point(colour = "blue", size = 3) +
  theme(text=element_text(size=10)) + 
  coord_flip() +
  labs(title = "The 10 happiest countries in 2018", x = "")
```

#### Top 10 happiest countries in 2019

```{r}
df1 %>%
  filter(Year == 2019) %>%
  arrange(-Happiness.Score) %>%
  slice_head(n=10) %>%
  ggplot(aes(reorder(Country, Happiness.Score), Happiness.Score)) +
  geom_point(colour = "blue", size = 3) +
  theme(text=element_text(size=10)) + 
  coord_flip() +
  labs(title = "The 10 happiest countries in 2019", x = "")
```


#### Top 10 happiest countries in 2020

```{r}
df1 %>%
  filter(Year == 2020) %>%
  arrange(-Happiness.Score) %>%
  slice_head(n=10) %>%
  ggplot(aes(reorder(Country, Happiness.Score), Happiness.Score)) +
  geom_point(colour = "blue", size = 3) +
  theme(text=element_text(size=10)) + 
  coord_flip() +
  labs(title = "The 10 happiest countries in 2020", x = "")
```


#### Top 10 happiest countries in 2021

```{r}
df1 %>%
  filter(Year == 2021) %>%
  arrange(-Happiness.Score) %>%
  slice_head(n=10) %>%
  ggplot(aes(reorder(Country, Happiness.Score), Happiness.Score)) +
  geom_point(colour = "blue", size = 3) +
  theme(text=element_text(size=10)) + 
  coord_flip() +
  labs(title = "The 10 happiest countries in 2021", x = "")
```


#### Top 10 happiest countries in 2022

```{r}
df1 %>%
  filter(Year == 2022) %>%
  arrange(-Happiness.Score) %>%
  slice_head(n=10) %>%
  ggplot(aes(reorder(Country, Happiness.Score), Happiness.Score)) +
  geom_point(colour = "blue", size = 3) +
  theme(text=element_text(size=10)) + 
  coord_flip() +
  labs(title = "The 10 happiest countries in 2022", x = "")
```

From the above graphs we can observe that,in 2015 Switzerland was the top happiest country. But it dropped to number two in 2016. Same as Denmark, which was the happiest country in 2016, but fell to number two in 2017. Norway was the happiest country in 2017. While Finland was the happiest country in 2018 and 2019.

### From 2015 to 2022, Mean Happiness score by regions:

```{r}
gg2 <- ggplot(df1 , aes(x = Region, y = Happiness.Score)) +
  geom_boxplot(aes(fill=Region)) + theme_bw() +
  theme(axis.text.x = element_text (angle = 90))

gg2
```


The top 3 happiness region are: Australia and New Zealand, North America and Western Europe.

### From 2015 to 2022, Mean happiness score by countries:

```{r}
df1 %>%
  group_by(Country) %>%
  summarise(mscore = mean(Happiness.Score)) %>%
  arrange(-mscore) %>%
  slice_head(n=10) %>%
  
  ggplot(aes(reorder(Country, mscore), mscore)) +
  geom_point() +
  theme_bw() +
  
  coord_flip() +
  labs(title = "Happiness Score by Country",
       x = "", y = "Average happiness score")
```
Top 3 happiness countries are: Denmark, Norway, Finland

### Top 10 Mean Happiness score by countries trends by years

```{r}
Top10_happy_country_DF = df1 %>%
  group_by(Country) %>%
  summarise(mscore = mean(Happiness.Score)) %>%
  arrange(-mscore) %>%
  slice_head(n=10)

Top10_happy_country_DF_list = c(Top10_happy_country_DF$Country)

df1_Top10_happy_country = subset(df1,Country %in% Top10_happy_country_DF_list)

ggplot(df1_Top10_happy_country,  aes(x = Year,y = Happiness.Score,color = Country))+  geom_line()
```

Only the happiness score of Finland is increasing dramatically from 2015-2022.

### Top 10 most progressive countries from 2015 - 2022:

```{r}
df1 %>%
  mutate(y = as.character(Year)) %>%
  select(y, Country, Region, Happiness.Score) %>%
  pivot_wider(names_from = y, values_from = Happiness.Score,
              names_prefix = "y_") %>%
  mutate(p = (y_2022 - y_2015)/y_2015 * 100) %>%
  arrange(-p) %>%
  slice_head(n = 10) %>%
  ggplot(aes(reorder(Country, p), p)) +
  geom_point() +
  theme_bw() +
  coord_flip() +
  labs(title = "The 10 most progressive countries from 2015 - 2022",
       y = "Percentage Increase of Happiness Score", x = "")
```

```{r}
Top10_Progress_country_df = df1 %>%
  mutate(y = as.character(Year)) %>%
  select(y, Country, Region, Happiness.Score) %>%
  pivot_wider(names_from = y, values_from = Happiness.Score,
              names_prefix = "y_") %>%
  mutate(p = (y_2019 - y_2015)/y_2015 * 100) %>%
  arrange(-p) %>%
  slice_head(n = 10)

Top10_Progress_country_df_list = c(Top10_Progress_country_df$Country)

df1_Top10_Progress_country = subset(df1,Country %in% Top10_Progress_country_df_list)

ggplot(df1_Top10_Progress_country,  aes(x = Year,y = Happiness.Score,color = Country))+  geom_line()
```

## Explore data by factors

```{r}
colnames(df1)
```
```{r}
head(df1)
```

### The mean value of the factors

```{r}
df1 %>%
  summarise(gdp = mean(Economy..GDP.per.Capita.),
            family = mean(Family),
            life.expectancy = mean(Health..Life.Expectancy.),
            freedom = mean(Freedom),
            generosity = mean(Generosity),
            corruption = mean(Trust..Government.Corruption.)) %>%
  pivot_longer(c(gdp, family, life.expectancy,freedom,generosity, corruption),
               names_to = "f", values_to = "value") %>%
  ggplot(aes(reorder(f, value), value)) +
  geom_bar(stat = "identity", fill = "orange", width = 0.55, alpha = 0.7) +
  geom_text(aes(label = paste0(round(value, 2)), vjust = -0.5)) +
  theme_bw() +
  labs(title = "The mean value of the factors" , y = "", x = "")
```
The family factor has the highest mean value, which is 1.09.

### Average value of happiness variables for different regions

```{r}
Happiness.Continent <- df1 %>%
                          select(-c(Year,Happiness.Rank))%>%
                          group_by(Region) %>%
                          summarise_at(vars(-Country), funs(mean(., na.rm=TRUE)))


Happiness.Continent.melt <- melt(Happiness.Continent)


# Faceting
ggplot(Happiness.Continent.melt, aes(y=value, x=Region, color=Region, fill=Region)) + 
  geom_bar( stat="identity") +    
  facet_wrap(~variable) + theme_bw() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  labs(title = "Average value of happiness variables for different regions", 
       y = "Average value") 
```

## Find Relationship using Scatter Plot of Happiness Score with each variable (include regression line)

### Scatter plot of Happiness Score with Economy_GDP per Capita (overall and by region)

```{r}
ggline1 = ggplot(df1, aes(x = Economy..GDP.per.Capita., y = Happiness.Score)) + 
  geom_point(size = .5, alpha = 0.8) +  
  geom_smooth(method = "lm", fullrange = TRUE) +
  theme_bw() + labs(title = "Scatter plot with regression line")

ggline1a = ggplot(df1, aes(x = Economy..GDP.per.Capita., y = Happiness.Score)) + 
  geom_point(aes(color=Region), size = .5, alpha = 0.8) +  
  geom_smooth(aes(color = Region, fill = Region), 
              method = "lm", fullrange = TRUE) +
  facet_wrap(~Region) +
  theme_bw() + labs(title = "Scatter plot with regression line")

ggline1
ggline1a
```

### Scatter plot of Happiness Score with Family (overall and by region)

```{r}
ggline2 = ggplot(df1, aes(x = Family, y = Happiness.Score)) + 
  geom_point(size = .5, alpha = 0.8) +  
  geom_smooth(method = "lm", fullrange = TRUE) +
  theme_bw() + labs(title = "Scatter plot with regression line")

ggline2a = ggplot(df1, aes(x = Family, y = Happiness.Score)) + 
  geom_point(aes(color=Region), size = .5, alpha = 0.8) +  
  geom_smooth(aes(color = Region, fill = Region), 
              method = "lm", fullrange = TRUE) +
  facet_wrap(~Region) +
  theme_bw() + labs(title = "Scatter plot with regression line")

ggline2
ggline2a
```

### Scatter plot of Happiness Score with Health_Life Expentancy (overall and by region)

```{r}
ggline3 = ggplot(df1, aes(x = Health..Life.Expectancy., y = Happiness.Score)) + 
  geom_point(size = .5, alpha = 0.8) +  
  geom_smooth(method = "lm", fullrange = TRUE) +
  theme_bw() + labs(title = "Scatter plot with regression line")

ggline3a = ggplot(df1, aes(x = Health..Life.Expectancy., y = Happiness.Score)) + 
  geom_point(aes(color=Region), size = .5, alpha = 0.8) +  
  geom_smooth(aes(color = Region, fill = Region), 
              method = "lm", fullrange = TRUE) +
  facet_wrap(~Region) +
  theme_bw() + labs(title = "Scatter plot with regression line")

ggline3
ggline3a
```

### Scatter plot of Happiness Score with Freedom (overall and by region)

```{r}
ggline4 = ggplot(df1, aes(x =Freedom, y = Happiness.Score)) + 
  geom_point(size = .5, alpha = 0.8) +  
  geom_smooth(method = "lm", fullrange = TRUE) +
  theme_bw() + labs(title = "Scatter plot with regression line")

ggline4a = ggplot(df1, aes(x =Freedom, y = Happiness.Score)) + 
  geom_point(aes(color=Region), size = .5, alpha = 0.8) +  
  geom_smooth(aes(color = Region, fill = Region), 
              method = "lm", fullrange = TRUE) +
  facet_wrap(~Region) +
  theme_bw() + labs(title = "Scatter plot with regression line")

ggline4
ggline4a
```

### Scatter plot of Happiness Score with Trust_Government Corruption (overall and by region)

```{r}
ggline5 = ggplot(df1, aes(x = Trust..Government.Corruption., y = Happiness.Score)) + 
  geom_point(size = .5, alpha = 0.8) +  
  geom_smooth(method = "lm", fullrange = TRUE) +
  theme_bw() + labs(title = "Scatter plot with regression line")

ggline5a = ggplot(df1, aes(x = Trust..Government.Corruption., y = Happiness.Score)) + 
  geom_point(aes(color=Region), size = .5, alpha = 0.8) +  
  geom_smooth(aes(color = Region, fill = Region), 
              method = "lm", fullrange = TRUE) +
  facet_wrap(~Region) +
  theme_bw() + labs(title = "Scatter plot with regression line")

ggline5
ggline5a
```

### Scatter plot of Happiness Score with Generosity (overall and by region)

```{r}
ggline6 = ggplot(df1, aes(x = Generosity, y = Happiness.Score)) + 
  geom_point(size = .5, alpha = 0.8) +  
  geom_smooth(method = "lm", fullrange = TRUE) +
  theme_bw() + labs(title = "Scatter plot with regression line")

ggline6a = ggplot(df1, aes(x = Generosity, y = Happiness.Score)) + 
  geom_point(aes(color=Region), size = .5, alpha = 0.8) +  
  geom_smooth(aes(color = Region, fill = Region), 
              method = "lm", fullrange = TRUE) +
  facet_wrap(~Region) +
  theme_bw() + labs(title = "Scatter plot with regression line")

ggline6
ggline6a
```

## Find Correlation using Correlation Matrix Heatmap

### Drop columns based on Heatmap Correlation

We should drop Year,Country,Happiness.Rank,Region column before computing the heatmap.

```{r}
dataset = select(df1,-c("Year","Country","Happiness.Rank","Region"))
head(dataset)
```

### Computing heatmap correlation

```{r}
library(corrplot)
Num.cols <- sapply(dataset, is.numeric)
Cor.data <- cor(dataset[, Num.cols])

corrplot(Cor.data, method = 'color') 
```

```{r}
library(GGally)

ggcorr(dataset, label = TRUE, label_round = 2, label_size = 3.5, size = 2, hjust = .85) +
  ggtitle("Correlation Heatmap") +
  theme(plot.title = element_text(hjust = 0.5))
```

## Categorize Happiness score into 3 level for classfication Algorithm:
* High,mid,low
* Adding new column Happy.Level into dataset
```{r}
rge_dif=round((max(dataset$Happiness.Score)-min(dataset$Happiness.Score))/3,3)

low=min(dataset$Happiness.Score)+rge_dif
mid=low+rge_dif

print(paste("range difference in happiness score: ",rge_dif))
print(paste('upper bound of Low grp',low))
print(paste('upper bound of Mid grp',mid))
print(paste('upper bound of High grp','max:',max(dataset$Happiness.Score)))
```
Transform “hapiness.Score” column into “Happy.Level” column

```{r}
dataset_level <- dataset %>%
  mutate(Happy.Level=case_when(
    Happiness.Score <=low  ~ "Low",
    Happiness.Score>low & Happiness.Score <=mid ~ "Mid",
    Happiness.Score >mid ~ "High"
  ))  %>%
  mutate(Happy.Level=factor(Happy.Level, levels=c("High", "Mid", "Low"))) %>%
  select(-Happiness.Score)
```


# Regression Models
## Split into train set (80%) and test set (20%)
```{r}
# Splitting the dataset into the Training set and Test set
set.seed(123) 
split=0.80
trainIndex <- createDataPartition(dataset$Happiness.Score, p=split, list=FALSE) 
data_train <- dataset[ trainIndex,] 
data_test <- dataset[-trainIndex,]

```

## Multiple Linear Regression for Happiness Score Prediction
### Train Multiple Linear Regression model with data_train
```{r}
# Fitting Multiple Linear Regression to the Training set
lm_model = lm(formula = Happiness.Score ~ .,
               data = data_train)
summary(lm_model)
```
An (adjusted) R2 that is close to 1 indicates that a large proportion of the variability in the outcome has been explained by the regression model.

A number near 0 indicates that the regression model did not explain much of the variability in the outcome.

Our adjusted R2 is 0.7697, which is good.

### Predict happiness score with data_test
```{r}
y_pred_lm = predict(lm_model, newdata = data_test)
Actual_lm = data_test$Happiness.Score

Pred_Actual_lm <- as.data.frame(cbind(Prediction = y_pred_lm, Actual = Actual_lm))

gg.lm <- ggplot(Pred_Actual_lm, aes(Actual, Prediction )) +
  geom_point() + theme_bw() + geom_abline() +
  labs(title = "Multiple Linear Regression", x = "Actual happiness score",
       y = "Predicted happiness score") +
  theme(plot.title = element_text(face="bold", size = (15)), 
        axis.title = element_text(size = (10)))
gg.lm
```

```{r}
data.frame(
  R2 = R2(y_pred_lm, data_test$Happiness.Score),
  RMSE = RMSE(y_pred_lm, data_test$Happiness.Score),
  MAE = MAE(y_pred_lm, data_test$Happiness.Score)
)
```

# Support Vector Regression for Happyniess Score Prediction
### Train SVR model with data_train
```{r}
library(e1071)

regressor_svr = svm(formula = Happiness.Score ~ .,
                data = data_train,
                type = 'eps-regression',
                kernel = 'radial')
```

## Predict happiness score with data_test

```{r}
# Predicting happiness score with SVR model
y_pred_svr = predict(regressor_svr,  newdata = data_test)

Pred_Actual_svr <- as.data.frame(cbind(Prediction = y_pred_svr, Actual = data_test$Happiness.Score))


Pred_Actual_lm.versus.svr <- cbind(Prediction.lm = y_pred_lm, Prediction.svr = y_pred_svr, Actual = data_test$Happiness.Score)


gg.svr <- ggplot(Pred_Actual_svr, aes(Actual, Prediction )) +
  geom_point() + theme_bw() + geom_abline() +
  labs(title = "SVR", x = "Actual happiness score",
       y = "Predicted happiness score") +
  theme(plot.title = element_text(face = "bold", size = (15)), 
        axis.title = element_text(size = (10)))
gg.svr
```

```{r}
data.frame(
  R2 = R2(y_pred_svr, data_test$Happiness.Score),
  RMSE = RMSE(y_pred_svr, data_test$Happiness.Score),
  MAE = MAE(y_pred_svr, data_test$Happiness.Score)
)
```

## Decision Tree Regression for Happiness Score Prediction
### Train Decision Tree Regression model with data_train

```{r}
library(rpart)
regressor_dt = rpart(formula = Happiness.Score ~ .,
                  data = data_train,
                  control = rpart.control(minsplit = 10))
```

## Predict happiness score with data_test
```{r}
# Predicting happiness score with Decision Tree Regression
y_pred_dt = predict(regressor_dt, newdata = data_test)

Pred_Actual_dt <- as.data.frame(cbind(Prediction = y_pred_dt, Actual = data_test$Happiness.Score))


gg.dt <- ggplot(Pred_Actual_dt, aes(Actual, Prediction )) +
  geom_point() + theme_bw() + geom_abline() +
  labs(title = "Decision Tree Regression", x = "Actual happiness score",
       y = "Predicted happiness score") +
  theme(plot.title = element_text(face = "bold", size = (15)), 
        axis.title = element_text(size = (10)))
gg.dt
```

```{r}
library(rpart.plot)
prp(regressor_dt)
```
```{r}
data.frame(
  R2 = R2(y_pred_dt, data_test$Happiness.Score),
  RMSE = RMSE(y_pred_dt, data_test$Happiness.Score),
  MAE = MAE(y_pred_dt, data_test$Happiness.Score)
)
```


# Classification Models
```{r}
# Splitting the dataset into the Training set and Test set
set.seed(123) 

split=0.80
trainIndex <- createDataPartition(dataset_level$Happy.Level, p=split, list=FALSE) 
data_train <- dataset_level[ trainIndex,] 
data_test <- dataset_level[-trainIndex,]
```

# cross validation
```{r}
tc <- trainControl(method = "repeatedcv", 
                   number=10,#10-fold cross validation 
                   classProbs = TRUE,
                   savePredictions = TRUE, 
                   repeats = 3,
                   ## Estimate class probabilities
                   summaryFunction = multiClassSummary,)
```

## K-Nearest Neighbor Classifier for predicting Happy Level
### Train K-Nearest Neighbours model with data_train
```{r}
set.seed(123)
library("MLmetrics")
model_knn <- train(
  Happy.Level~., 
  data=data_train, 
  trControl=tc,
  preProcess = c("center","scale"),
  method="knn",
  metric='Accuracy',
  tuneLength=20
  ) 

model_knn

```
```{r}
plot(model_knn)
```

### Predict happiness level by K-Nearest Neighbours model

```{r}

pred_knn <- predict(model_knn, data_test)

cm_knn<-confusionMatrix(pred_knn, data_test$Happy.Level)

cm_knn
```

### Feature Importance
```{r}
# Create object of importance of our variables 
knn_importance <- varImp(model_knn) 

# Create box plot of importance of variables
ggplot(data = knn_importance, mapping = aes(x = knn_importance[,1])) + # Data & mapping
  geom_boxplot() + # Create box plot
  labs(title = "Variable importance: K-Nearest Neighbours ") + # Title
  theme_light() # Theme
```
## Naive Bayes Classification model for predicting Happy Level
### Train Naive Bayes Classification model model with data_train
```{r}
model_nb <- train(Happy.Level~.,
                  data_train,
                  method="naive_bayes",
                  preProcess = c("center","scale"),
                  metric='Accuracy',
                  trControl=tc)

model_nb
```
```{r}
plot(model_nb)
```
### Predict happiness Level by Naive Bayes
```{r}
pred_nb <- predict(model_nb, data_test)

cm_nb<-confusionMatrix(pred_nb, data_test$Happy.Level)

cm_nb
```
## Feature Importance

```{r}
# Create object of importance of our variables 
nb_importance <- varImp(model_nb) 

# Create box plot of importance of variables
ggplot(data = nb_importance, mapping = aes(x = nb_importance[,1])) + # Data & mapping
  geom_boxplot() + # Create box plot
  labs(title = "Variable importance: Naive Bayes model") + # Title
  theme_light() # Theme
```
### Model Evaluation by AUC

```{r}
model_list <- list(KNN = model_knn, NB=model_nb)
resamples <- resamples(model_list)

bwplot(resamples, metric="AUC")
```

### Model Evaluation by Prediction Accurancy

```{r}
data.frame(
  K_Nearest_Neighbours= cm_knn$overall[1],
  Naive_Bayes=  cm_nb$overall[1]
)
```
